---
title: "Hobby Airport - EDA"
author: "RavMin"
date: "March 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Houston Hobby Airport data

#For this month's data jam meet up, we worked with data from Monday at the Houston Hobby airport. This data is sourced from FlightAware's API.

#We have the data available in csv, each in their respective directories.

```{r}

make_csv_url <- function(name){
  url <- paste('https://raw.githubusercontent.com/houstondatavis/data-jam-february-2017/data-pipeline/', name, '.csv', sep='')

  return(url)
}

flights <- read.csv(make_csv_url('flights'))

routes <- read.csv(make_csv_url('routes'))

tracks <- read.csv(make_csv_url('tracks'))

weather <- read.csv(make_csv_url('weather'))

```

# Checking the unique flight$ident in flights data
```{r}
dim(flights)
flight_numbers <- unique(flights$ident, incomparables = FALSE)

```

# Extract the flight ident from the routes flight_id
```{r}
routes$flight_no <- substring(routes$flight_id, 1,7)
head(routes)

```
# Extract flight ident no from the tracks flight_id

```{r}
tracks$flight_no <- substring(tracks$flight_id, 1,7)
head(tracks)


```

```{r}
library(leaflet)
library(maps)
map_usa = map("state", fill = TRUE, plot = FALSE)
m = leaflet(map_usa) %>% addTiles() %>%
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE)
  
m

```


##Plotting the route plan of flights into Houston 
```{r}

m %>%
  addPolylines(data = routes, lat = ~ latitude, lng = ~ longitude, color="red", , weight = 1)%>% 
  setView(-95.27889, 29.64542, zoom = 6)%>%
  addMarkers(-95.27889, 29.64542, popup = "Destination Houston")

```

##Plotting the Tracks (Flight-specific location tracking data) for all flights into Houston 

```{r}

m %>%
  addPolylines(data = tracks, lat = ~ latitude, lng = ~ longitude, color="red", , weight = 1)%>% 
  setView(-95.27889, 29.64542, zoom = 6)%>%
  addMarkers(-95.27889, 29.64542, popup = "Destination Houston")

```

### We can see that flight tracks ar too different from their planned routes. Lets look deeper into a few individual flights

## Plots for flight SWA4569
```{r}

route_SWA4569 <- subset(routes, flight_no == 'SWA4569')
track_SWA4569 <- subset(tracks, flight_no == 'SWA4569')


m %>%
  addPolylines(data = route_SWA4569, lat = ~ latitude, lng = ~ longitude, color="green", weight = 4)%>%   addPolylines(data = track_SWA4569, lat = ~ latitude, lng = ~ longitude, color="blue", weight = 4)%>% 
  setView(-95.27889, 29.64542, zoom = 6)%>%
  addMarkers(-95.27889, 29.64542, popup = "Destination Houston")

```

## Plots for flight SWA3020

```{r}
route_SWA3020 <- subset(routes, flight_no == 'SWA3020')
track_SWA3020 <- subset(tracks, flight_no == 'SWA3020')

for(row in 1:length(route_SWA3020)){
  if (route_SWA3020$order[row] == 0){
    origin_lat = route_SWA3020$latitude[row]
    origin_long = route_SWA3020$longitude[row]
  }
}
origin_lat
origin_long


```

```{r}
m %>%
  addPolylines(data = route_SWA3020, lat = ~ latitude, lng = ~ longitude, color="green", weight = 4)%>%   addPolylines(data = track_SWA3020, lat = ~ latitude, lng = ~ longitude, color="blue", weight = 4)%>% 
  setView(-95.27889, 29.64542, zoom = 8)%>%
  addMarkers(-86.67817, 36.12447, popup = "Origin Airport")%>%
  addMarkers(-95.27889, 29.64542, popup = "Destination Houston") 

```

Lets look at flight SWA1775
```{r}
route_SWA1775 <- subset(routes, flight_no == 'SWA1775')
track_SWA1775 <- subset(tracks, flight_no == 'SWA1775')

for(row in 1:length(route_SWA1775)){
  if (route_SWA1775$order[row] == 0){
    origin_lat = route_SWA1775$latitude[row]
    origin_long = route_SWA1775$longitude[row]
  }
}
origin_lat
origin_long




```
```{r}
m %>%
  addPolylines(data = route_SWA1775, lat = ~ latitude, lng = ~ longitude, color="green", weight = 4)%>%   addPolylines(data = track_SWA1775, lat = ~ latitude, lng = ~ longitude, color="blue", weight = 4)%>% 
  setView(-95.27889, 29.64542, zoom = 8)%>%
  addMarkers(-87.75242, 41.78597, popup = "Origin Airport")%>%
  addMarkers(-95.27889, 29.64542, popup = "Destination Houston") 

```


## Convert datetime form UTC
```{r}
convertdatetime <- function(unix_time){
  date1 <- as.POSIXct(unix_time, origin = "1970-01-01")
  #return(time)
}

strip_time <- function(date1){
  time <- strftime(date1, format = '%H:%M:%S')

}

tracks["new_date"] <- lapply(tracks["timestamp"], convertdatetime)
tracks["new_time"] <- lapply(tracks["new_date"], strip_time)
View(tracks)



weather["new_date"] <- lapply(weather["time"], convertdatetime)
weather["new_time"] <- lapply(weather["new_date"], strip_time)
View(weather)
```


## Lets plot the altitude of flight SWA4569 against the time
```{r}
library(plotly)
flight_traj1 <- plot_ly(data = track_SWA4569, x = ~longitude, y = ~latitude, z=~altitude, type = 'scatter3d', mode = 'lines', opacity = 1, line = list(width = 2, color = 'red', reverscale = FALSE))

flight_traj1



```

```{r}
flight_traj2 <- plot_ly(data = track_SWA3020, x = ~longitude, y = ~latitude, z=~altitude, type = 'scatter3d', mode = 'lines', opacity = 1, line = list(width = 2, color = 'red', reverscale = FALSE))

flight_traj2

```


```{r}
flight_traj3 <- plot_ly(data = track_SWA1775, x = ~longitude, y = ~latitude, z=~altitude, type = 'scatter3d', mode = 'lines', opacity = 1, line = list(width = 2, color = 'red', reverscale = FALSE))

flight_traj3



```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
